Understanding the impact of electronically-delivered parenting advice

### 2019-02-28 paper/cogsci2019/cogsci2019.tex
working on E1 results and intro

\author{{\large\bf Hanwen Vivian~Zhang} (\texttt{vivian3@stanford.edu}) \\ {\large\bf George~Kachergis} (\texttt{george.kachergis@gmail.com}) \\ {\large\bf Michael C.~Frank} (\texttt{mcfrank@stanford.edu}) \\  Department of Psychology, Stanford University \\  Palo Alto, CA 94301 USA}

@@

\begin{abstract}
Early parenting practices play an important role in shaping the future
outcomes of young children (Hart \& Risley, 1995; Heckman, 2006). In
particular, high quality early interactions and early language input
appear to facilitate more effective language learning and higher levels
of school performance. The rise in electronic parenting applications
(``apps'') holds the promise of delivering low-cost, positive
interventions on parenting style to a variety of different populations.
Of special interest are the parents of very young children, who are
often difficult to reach in other ways. Yet little is known about the
effects of communicating to parents through app-based interventions.
MTLD, types and tokens, joint attention, proportion of
nouns/verbs/adjectives, affect, concreteness, mean length of
utterance\ldots{}

@@

Children spend a large portion of their waking time at play in their
environment. Playing with objects may allow them to discover hidden
features, as well as to build a causal understanding of how the objects
interact. Beyond learning object features, the laws of physics, and
causal relationships (e.g., Schulz \& Bonawitz, 2007), during play
children practice motor skills that help them effectively navigate their
world (Singer, Golinkoff, \& Hirsh-Pasek, 2006). Social play can help
children learn about human relationships, both through imitation of
adult behaviors and by experiencing and learning to process emotional
events (Singer et al., 2006).

@@

Children's early play behaviors are often assisted by more skilled and
knowledgeable play partners such as their caregivers and older siblings
(Kaye, 1970). Under such expert guidance, children are encouraged and
motivated to engage in more advanced play, undertaking explorations that
push the boundaries of what they would be able to do unaided (Vygotsky,
1980). These tutorial interactions have been shown to be important
components of child development (Wood, Bruner, \& Ross, 1976). Thus,
with the knowledge of both play and tutorial interactions, guided play,
which consists of both active and enjoyable activities as well as close
guidance of adults (Hirsh-Pasek \& Golinkoff, 2008) has drawn
researchers' interest. A study on preschoolers by Weisberg showed that
guided play scaffolds the environment while still allowing children to
maintain a large degree of control, and it outperforms
direct-instruction approaches in encouraging a variety of positive
academic outcomes (Weisberg, Hirsh-Pasek, \& Golinkoff, 2013). Another
study by Massey specifically found that guided play could facilitate
children's vocabulary and comprehensive language development and
subsequent literacy skills (Massey, 2013).

@@

Thus, including guided play in parenting practices early on is crucial
and will help with children's later performances in academic and social
situations. To intervene in parenting practices, Suskind and others did
a parent-directed home-visiting intervention experiment in 2015 for
children of low socioeconomic status. They showed that parents in
experimental group have greater knowledge of language development and
this effect sustained four months after the intervention. However, more
interactive outcomes, like adult word tokens, conversational turn
counts, and child vocalization counts, though increased significantly
during the experiment, did not sustain after the intervention (D. L.
Suskind et al., 2015). That changes did not sustain could be due to the
intervention itself, or could merely be that the methods of
home-visiting is not sustainable enough for parents to easily and
constantly get parenting advice. Thus, new methods of delivering
parenting advice should be considered.

@@

With the development of technology, digitally-delivered methods could
address some logistic barriers to face-to-face delivery methods.
Breitenstein has reviewed 11 studies on them and showed good promise for
the efficacy of interventions using digitally-delivered methods
(Breitenstein, Gross, \& Christophersen, 2014). However, the parent and
child outcomes assessed in the review (e.g.infant positive behaviors,
satisfaction, emotional symptoms etc.) did not address the nature or
quality of parent-infant interactions at a detailed level, like if the
interventions lead to children paying more attention, or vocabulary
changes in parents' language usages. Although digitally-delivered
ac-tivities are designed to promote learning and cognitive devel-opment,
it is unclear how they might affect these dimensions of parent-child
interactions. Thus, we want to conduct this experiment to explore if and
how digital scaffolding of activities affect the social and linguistic
characteristics of parent-child ineractions. The quality of parent-child
interactions can be measured by both the social engagement of parents
(e.g., joint attention to objects in the environment)(Bigelow, MacLean,
\& Proctor, 2004) and the quality of language (e.g., vocabulary
diversity) (Malvern, Richards, Chipere, \& Durán, 2004).

@@

\begin{CodeChunk}
\begin{CodeOutput}
[1] 45
\end{CodeOutput}
\end{CodeChunk}

@@

60 infants (F = 42, M = 18) aged 6-24 months (20 6-11.9 month-olds, 20
12-17.9 month-olds, and 20 18-24 month-olds) and their parents
participated in a museum in northern California. We included infants who
were exposed to English at least 50 percent of the time (n = 58) or who
were exposed less but whose participating parent reported that they
primarily speak English with their child at home (n = 2). Sixty-one\% of
participants (n = 37) had been exposed to 2 or more languages as
indicated by their parent. Parents identified their children as White (n
= 25), Asian (n = 11), African American/Black (n = 2), Biracial (n =
12), other (n = 5), or declined to state (n = 5). Fifteen parents
reported their child was of Hispanic origin. Parents tended to be highly
educated, with a highest level of education reported as high school (n =
5), some college (n = 6), four-year college (n = 14), some graduate
school (n = 1), completed graduate school (n = 28) or declined to state
(n = 6).

@@

Stimuli included videos from a commercial parenting application, Kinedu
Inc. The videos were designed to show activities parents could perform
with their child to foster cognitive and physical development, and were
targeted to the child's age and development. In each video, an adult and
child perform the activity while a narrator explains the activity and
its purpose. We selected two videos for each of three age groups in our
sample (6-11.9 months, 12-17.9 months, 18-23.94 months). More
information about the specific videos is available in the Appendix.
Participants were also given a set of props corresponding to those in
the video they watched, so that they could complete the activity. The
props associated with each video are listed in the Appendix.

Participants were randomly assigned to either the \emph{Video} condition
or the \emph{Control} condition. Participants in the Activity Video
condition were assigned to one of the two videos available for their
child's age group. The No-Video Control condition was yoked to the
Activity Video condition such that for every participant in the Activity
Video condition who saw a particular video and received the associated
props, a participant in the No-Video Control condition received the same
props but did not watch the corresponding video. Parents also completed
the Parenting Attitudes Questionnaire (PAQ). The PAQ measures parents'
attitudes about parenting and child development along three dimensions:
rules and respect, early learning, and affection and attachment.

@@

After providing informed consent, parents in the Activity Video
condition watched the assigned video on a laptop with headphones. To
ensure that parents could give the video their full attention, the
experimenter played with the infant with a set of toys (different from
the experimental props used in the study) while the video was being
played. Immediately following the video, each parent-child dyad was
provided with the props to complete the activity they had viewed. The
toys were placed on a large foam mat, and parents were instructed to sit
on the mat with their child and re-create the activity they had viewed
for a period of three minutes. In the No-Video control condition, after
consenting parents were told to play with their child as they would at
home with the provided props for a period of three minutes. They were
not given any additional instructions about how to use the props.

In both conditions, two video cameras were used to record the play
session from different angles, and parents were fitted with a wireless
audio recorder to record their speech to their child. After three
minutes of play had elapsed, parents were told they could stop playing
and cameras and audio were turned off. Parents were then asked to
complete the PAQ before being debriefed.

@@

The video of each session was manually coded for episodes of joint
attention using the Datavyu software. The video taken at floor level was
coded by default, but the other video was referred to if the
participants were not visible or if there was a technological error with
the first camera. For each participant we coded for episodes of
coordinated joint attention, episodes of passive joint attention, and
parental bids for joint attention. We defined parental bids for joint
attention as any attempt to initiate joint attention (i.e label, point,
or otherwise draw attention to an object) that does not result in
passive or coordinated joint attention. If more than 3 seconds elapsed
between bids, they were coded as separate attempts. An episode of joint
attention was considered passive if both participants visually focused
on an object for a minimum of 3 seconds but the child did not
acknowledge the parent. If either participant looked away from the
object for less than 3 seconds and then returned to the same object it
was considered part of the same period of joint attention. Episodes
joint attention were considered coordinated if both participants
visually focused on an object for a minimum of 3 seconds and at some
point in the interaction the child indicated awareness of interaction
with some overt behavior toward the parent. This could be looks to the
parent's face, gestures, vocalizations, or turn-taking. If either
participant looked away for less than 3 seconds and then returned to the
same object, it was coded as part of the same period of joint attention.
A second coder independently coded a third of the videos (i.e., 20 of
the 60 videos, approximately equally distributed across ages) to
establish reliability. The two coders achieved reliability with ICC =
0.80 with 95\% confident interval = 0.57-0.92 (p \textless{} 0.05) for
number of parent bids; ICC = 0.20 with 95\% confident interval =
-0.26-0.58 for number of passive joint attention established; ICC = 0.66
with 95\% confident interval = 0.32-0.85 (p \textless{} 0.05) for number
of coordinated joint attention established; ICC = 0.24 with 95\%
confident interval = -0.21-0.61 for length of passive joint attention,
and ICC = 0.62 with 95\% confident interval = 0.27-0.83 (p \textless{}
0.05) for length of passive joint attention.

@@

The data was analyzed according to our preregistration:
\href{https://osf.io/2bpdf/}{link}.

@@

We first calculated Type/Token Ratio (TTR) for each transcript. ..

Fig 1 - TTR, maybe also Total Tokens and Total Types (if subfigures fit)
(otherwise a table with mean and SD of tokens, types, and TTR) However,
note that TTR is a measure that is correlated with the length of a text.
\ldots{} MTLD is a better measure of lexical diversity (McCarthy \&
Jarvis, 2010).

@@

\caption[Mean lexical diversity scores by condition (left]{Mean lexical diversity scores by condition (left: Type/Token ratio, right: MTLD) in Experiment 1.}\label{fig:e1lex_div}
\end{figure}
\end{CodeChunk}

@@

\subsection{Footnotes}\label{footnotes}

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. You can also use markdown formatting to include
footnotes using this syntax.\footnote{Sample of a markdown footnote.}

@@

\hypertarget{refs}{}
\hypertarget{ref-Bigelow2004}{}
Bigelow, A. E., MacLean, K., \& Proctor, J. (2004). The role of joint
attention in the development of infants' play with objects.
\emph{Developmental Science}, \emph{7}, 518--526.

@@

@@

>>>>

\author{{\large\bf Hanwen Vivian~Zhang} (\texttt{vivian3@stanford.edu}) \\ {\large\bf George~Kachergis} (\texttt{george.kachergis@gmail.com}) \\ {\large\bf Michael C.~Frank} (\texttt{mcfrank@stanford.edu}) \\  Department of Psychology, Stanford University \\  Stanford, CA 94305 USA}

@@

\begin{abstract}
Early parenting practices play an important role in shaping the future
outcomes of young children (Hart \& Risley, 1995; Heckman, 2006). In
particular, high-quality early interactions and language input appear to
facilitate language learning and result in higher levels of school
performance. The rise of phone- and tablet-based parenting applications
(``apps'') holds the promise of delivering low-cost, positive
interventions on parenting style to a variety of different populations.
Of special interest are the parents of very young children, who are
often difficult to reach in other ways. Yet little is known about the
effects of communicating to parents through app-based interventions. In
a study of one commercial app offering a collection of age-appropriate
activity videos, we find that the quality of parent-child interactions
increases in some ways as a result of using the app. Specifically, the
lexical diversity of parents' child-directed speech increases, and
measures of joint attention show\ldots{}

@@

Young children spend a large portion of their waking time at play,
variously manipulating objects, exploring their environment, and
interacting with caregivers and peers. Playing with objects allows them
to discover hidden object properties and relations, and to build a
causal understanding of how objects interact (e.g., Schulz \& Bonawitz,
2007). Meanwhile, play also gives children an opportunity to set and
achieve goals (e.g., build a tower) and to practice a wide range of
motor skills (e.g., stacking) that will help them navigate the world
(Singer, Golinkoff, \& Hirsh-Pasek, 2006). Social play can help children
learn about human relationships, both through imitation of adult
behaviors and by experiencing and learning to process emotional events
such as failures (Singer et al., 2006). Of course, young children are
rarely playing in isolation: caregivers often provide encouragement and
guidance while scaffolding a child's play (Kaye, 1970; Wood, Bruner, \&
Ross, 1976). Early parenting practices have been shown to play an
important role in shaping the future outcomes of young children (Hart \&
Risley, 1995; Heckman, 2006). While many interventions have trouble
reaching the parents of very young children, digital delivery of
parenting advice by way of

@@

Children's early play behaviors are often assisted by more skilled and
knowledgeable play partners such as their caregivers and older siblings
(Kaye, 1970). Under such expert guidance, children are encouraged and
motivated to engage in more advanced play, undertaking explorations that
push the boundaries of what they would be able to do unaided (Vygotsky,
1980). These tutorial interactions have been shown to be important
components of child development (Wood et al., 1976). Thus, with the
knowledge of both play and tutorial interactions, guided play, which
consists of both active and enjoyable activities as well as close
guidance of adults (Hirsh-Pasek \& Golinkoff, 2008) has drawn
researchers' interest. A study of preschoolers showed that guided play
scaffolds the environment while still allowing children to maintain a
large degree of control, and it outperforms direct-instruction
approaches in encouraging a variety of positive academic outcomes
(Weisberg, Hirsh-Pasek, \& Golinkoff, 2013). Another study found that
guided play could facilitate children's vocabulary and comprehensive
language development and subsequent literacy skills (Massey, 2013).

@@

Thus, including guided play in parenting practices early on is crucial
and will help with children's later performances in academic and social
situations. To intervene in parenting practices, Suskind and others did
a parent-directed home-visiting intervention experiment in 2015 for
children of low socioeconomic status. They showed that parents in
experimental group have greater knowledge of language development and
this effect sustained four months after the intervention. However,
although a number of interactive measures increased during the
experiment, including the number of word tokens, conversational turns,
and child vocalizations, these increases did not sustain after the
intervention (D. L. Suskind et al., 2015). That changes did not sustain
could be due to the intervention itself, or could merely be that the
methods of home-visiting is not sustainable enough for parents to easily
and constantly get parenting advice. Thus, new methods of delivering
parenting advice should be considered.

@@

With the widespread use of smartphones and tablets worldwide,
digitally-delivered interventions could address many of the logistical
barriers that have limited scaling up face-to-face delivery methods. A
review of 11 studies verified that digital delivery is a promising means
of effectively distributing interventions (Breitenstein, Gross, \&
Christophersen, 2014). However, the parent and child outcomes assessed
in the review (e.g.infant positive behaviors, satisfaction, emotional
symptoms etc.) did not address the nature or quality of parent-infant
interactions at a detailed level, like if the interventions lead to
children paying more attention, or vocabulary changes in parents'
language usages. Although digitally-delivered ac-tivities are designed
to promote learning and cognitive development, it is unclear how they
might affect these dimensions of parent-child interactions. Thus, we
want to conduct this experiment to explore if and how digital
scaffolding of activities affect the social and linguistic
characteristics of parent-child ineractions. The quality of parent-child
interactions can be measured by both the social engagement of parents
(e.g., joint attention to objects in the environment)(Bigelow, MacLean,
\& Proctor, 2004) and the quality of language (e.g., vocabulary
diversity) (Malvern, Richards, Chipere, \& Durán, 2004).

@@

@@

60 infants (F = 42, M = 18) aged 6-24 months (20 6-11.9 month-olds, 20
12-17.9 month-olds, and 20 18-24 month-olds) and their parents
participated in a museum in northern California. We included infants who
were exposed to English at least 50 percent of the time (n = 58) or who
were exposed less but whose participating parent reported that they
primarily speak English with their child at home (n = 2). Sixty-one\% of
participants (n = 37) had been exposed to two or more languages as
indicated by their parent. Parents identified their children as White (n
= 25), Asian (n = 11), African American/Black (n = 2), Biracial (n =
12), other (n = 5), or declined to state (n = 5). Fifteen parents
reported their child was of Hispanic origin. Parents tended to be
highly-educated, with reports of highest level of education ranging from
completed high school (n = 5), some college (n = 6), four-year college
(n = 14), some graduate school (n = 1), to completed graduate school (n
= 28) or declined to state (n = 6).

@@

Stimuli included videos from the Kinedu (Kinedu Inc.) commercial
parenting application. The videos were designed to show activities to
parents that they could perform with their child in order to foster
cognitive and physical development, and were targeted to the child's age
and level of development. In each video, an adult and child perform the
activity while a narrator explains the activity and its purpose. We
selected two videos for each of three age groups in our sample (6-11.9
months, 12-17.9 months, 18-23.94 months). More information about the
specific videos is available in the Appendix. Participants were also
given a set of props corresponding to those in the video they watched so
that they could complete the activity. The props associated with each
video are listed in the Appendix.

Participants were randomly assigned to either the \emph{Video} condition
or the \emph{Control} condition. Participating parents in the Video
condition were assigned to watch one of the two activity videos
available for their child's age group. Participating parents in the
Control condition watched no activity video, and were merely asked to
play with their child as they normally would. The Control condition was
yoked to the Activity Video condition such that for every participant in
the Video condition who saw a particular video and received the
associated props, a participant in the Control condition received the
same props but did not watch the activity video. Parents also completed
the Parenting Attitudes Questionnaire (PAQ). The PAQ measures parents'
attitudes about parenting and child development along three dimensions:
rules and respect, early learning, and affection and attachment.

@@

After providing informed consent, parents in the Video condition watched
the assigned activity video on a laptop with headphones. To ensure that
parents could give the video their full attention, the experimenter
played with the infant with a set of toys (different from the
experimental props used in the study) while the video was being played.
Immediately following the video, each parent-child dyad was provided
with the props to complete the activity they had viewed. The toys were
placed on a large foam mat, and parents were instructed to sit on the
mat with their child and re-create the activity they had viewed for a
period of three minutes. In the Control condition, after consenting
parents were told to play with their child as they would at home with
the provided props for a period of three minutes. They were not given
any additional instructions about how to use the props.

In both conditions, two video cameras were used to record the play
session from different angles, and parents were fitted with a wireless
audio recorder to record their child-directed speech. After three
minutes of play had elapsed, parents were told they could stop playing
and cameras and audio were turned off. Parents were then asked to
complete the PAQ before being debriefed.

@@

The video of each session was manually coded for episodes of joint
attention using the Datavyu software (Team, 2014). The video taken at
floor level was coded by default, but the other video was referred to if
the participants were not visible or if there was technical difficulty
with the first camera. Each session's video was coded for episodes of
coordinated joint attention, episodes of passive joint attention, and
parental bids for joint attention. Parental bids for joint attention
were defined as any attempt to initiate joint attention (i.e labeling,
pointing, or otherwise drawing attention to an object) that did not
result in passive or coordinated joint attention. If more than 3 seconds
elapsed between bids, they were coded as separate attempts. An episode
of joint attention was considered passive if both participants visually
focused on an object for a minimum of 3 seconds but the child did not
acknowledge the parent. If either participant looked away from the
object for less than 3 seconds and then returned to the same object it
was considered part of the same period of joint attention. Episodes
joint attention were considered coordinated if both participants
visually focused on an object for a minimum of 3 seconds and at some
point in the interaction the child indicated awareness of interaction
with some overt behavior toward the parent. This could be looks to the
parent's face, gestures, vocalizations, or turn-taking. If either
participant looked away for less than 3 seconds and then returned to the
same object, it was coded as part of the same period of joint attention.
A second coder independently coded a third of the videos (i.e., 20 of
the 60 videos, approximately equally distributed across ages) to
establish reliability. The two coders had a reliability of ICC = 0.80
with 95\% confident interval (CI) = {[}0.57,0.92{]} (p \textless{} 0.05)
for number of parent bids for JA; ICC = 0.20 with 95\% CI =
{[}-0.26,0.58{]} for number of passive JA episodes; ICC = 0.66 with 95\%
CI = {[}0.32,0.85{]} (p \textless{} 0.05) for number of coordinated JA
episodes; ICC = 0.24 with 95\% CI = {[}-0.21,0.61{]} for total duration
of passive JA episodes, and ICC = 0.62 with 95\% CI = {[}0.27,0.83{]} (p
\textless{} 0.05) for total duration of passive (coordinated??) JA
episodes.

@@

The transcripts and hand-coded behavioral data was analyzed according to
our preregistration\footnote{Preregistration:
  \url{https://osf.io/2bpdf/}{]}}. Below we first describe the lexical
diversity results, followed by the joint attention results.

@@

Parents' child-directed speech during the play sessions was transcribed.
For each transcript, the words were lemmatized using Honnibal (2017),
and the word \emph{types} (unique words) and \emph{tokens} (total words)
were then tallied and the type-token ratio (TTR) calculated as a measure
of lexical diversity. Although TTR was our preregistered measure of
lexical diversity, TTR is correlated with the length of a text, whereas
the measure of textual lexical diversity (MTLD) is not (McCarthy \&
Jarvis, 2010). Thus, we also measure lexical diversity with MTLD, which
is calculated as the mean length of sequential word strings in a text
that maintain a given TTR value (here, .720).

We fit a mixed-effects linear regression predicting TTR as a function of
condition, age (scaled and 0-centered), gender, and parent's education
level with a random intercept per video using lme4 (Bates, Mächler,
Bolker, \& Walker, 2015). There was significantly lower TTR in the Video
condition (mean: 0.32) than in the Control condition (mean: 0.43,
\(\beta=-.12\), t(39.9) = 4.25, \emph{p}\textless{}.001). There were no
significant effects of age, gender, or parent's level of education. A
similar mixed-effects linear regression instead predicting MTLD also
found significantly lower lexical diversity in the Video condition
(mean: 15.7) than in the Control condition (mean: 21.9,
\(\beta=-11.67\), t(43) = 3.08, \emph{p}\textless{}.01), with no other
significant effects. Figure 1 shows the mean of each lexical diversity
measure (TTR and MTLD) by condition.

We also conducted similar regressions predicting the number of word
tokens and types, finding only a significant effect of condition on the
number of word tokens (\(\beta=57.23\), t(34.4)=2.19,
\emph{p}\textless{}.05), with parents using more words in the Video
condition (mean: 225, 95\% CI: {[}197,252{]}) than in the Control
condition (mean: 165, 95\% CI: {[}139,193{]}).

(Table with mean and SD of tokens, types, and TTR)

@@

\caption[Mean lexical diversity scores by condition (left]{Mean lexical diversity scores by condition (left: Type/Token ratio, right: MTLD) in Experiment 1. Error bars show bootstrapped 95 percent confidence intervals (CIs).}\label{fig:e1lex_div}
\end{figure}
\end{CodeChunk}

@@

@@

\hypertarget{refs}{}
\hypertarget{ref-lme4}{}
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting linear
mixed-effects models using lme4. \emph{Journal of Statistical Software},
\emph{67}(1), 1--48. \url{http://doi.org/10.18637/jss.v067.i01}

\hypertarget{ref-Bigelow2004}{}
Bigelow, A. E., MacLean, K., \& Proctor, J. (2004). The role of joint
attention in the development of infants' play with objects.
\emph{Developmental Science}, \emph{7}, 518--526.

@@

\hypertarget{ref-spacy2}{}
Honnibal, I., Matthew AND Montani. (2017). SpaCy 2: Natural language
understanding with bloom embeddings, convolutional neural networks and
incremental parsing. \emph{To Appear}.

@@

\hypertarget{ref-datavyu}{}
Team, D. (2014). Datavyu: A video coding tool. \emph{Databrary Project}.
Retrieved from \url{http://datavyu.org}
